---
title: "Race"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
source("../Race/demoChg.R", local = knitr::knit_global())

```
<center>
**THIS IS A DRAFT. DO NOT DISTRIBUTE. DO NOT CITE WITHOUT AUTHOR'S PERMISSION**
</center>

#
#


<a href="#racemap">Jump to the map</a>

Stakeholders also believe that Malheur is experiencing the “white flight” phenomenon, with middle-class, white families leaving the County for cities like Fruitland. While the data collected by the Census Bureau at the tract and county level does not allow us to see the movement of families between the two regions by race, we can follow aggregate changes in the racial makeup of the two counties between 2010 and 2018. We can look at changes in the white, non-Hispanic and non-white populations by examining the changes within each group and across all groups. This distinction is important because a population can become a smaller share of the overall population, but still see an increase in real numbers. This type of phenomenon will occur if one group grows at a faster rate than another. 

<font color="orange">Overall: </font>Malheur has a sizeable non-white population that has only grown between 2010 and 2018. Payette County has also seen an increase in its non-white population, though the white population is a significant majority in that County. While the total share of white individuals has gone down in both counties, the decrease was much larger in Malheur. 

In 2018, 61.5 percent (2018) of all those who lived in Malheur County were white, a slight decrease from 64 percent in 2010. Within the white population, there was a 7.3 percent decrease. In comparison, white individuals make up 82.2 percent of the state of Oregon (2018). The non-white population in the County increased by 5.2 percent, from 36 to 39 percent overall. The Latinx population increased from 30 percent in 2010 to 33.4 percent in 2018, a 7.1 percent increase within the population. The Black, Asian American and Pacific Islander (AAPI), and Native population are very small in the County, but also saw an increase. Unfortunately, further disaggregation of this data is not possible because of the significant measurement error present with such a small population.

While we cannot say with certainty that the area has experienced the white flight phenomenon, we do notice that a similar trend is present in Payette County. Like Malheur County, Payette County has seen a decrease in the white population (both within the communities and overall). Between 2010 and 2018, the white population in Payette County decreased from 82 percent to 78.5 percent. Within the white community, that was a decrease of 1.6 percent. This is a much smaller decrease than is observed in Malheur County. The Latinx population increased from 14.2 percent to 17.2 percent. The increase within the population was 24 percent, a much larger increase than the Latinx community in Malheur County. This increase is likely because Malheur originally had a much larger Latinx community. As was the case with Malheur County, the AAPI, Native, and Black populations are very small in Payette County. Overall, the nonwhite population saw an increase from 18 percent to 21 percent. 

Finally, the increase in the non-white population was large in the tract that includes Fruitland, while the share of white individuals did not decrease much. This suggests that much of the population growth in Fruitland was due to an increase in the non-white population. In the tract that includes Ontario, the rate of increase in the non-white population is similar to the rate of the decrease in the white population.  



<font color = "orange">Why can’t we make a definitive statement about white flight? </font> The publically available data does not allow us to observe individuals moving between the two regions. White families may have left Malheur County to move to Payette County. However, the data does not track individuals moving between the region at the tract or county level. This lack of granular data makes it difficult for researchers to observe the movement of families between the two areas. We recommend the Board conduct a survey if this is a phenomenon that they are interested in further understanding. 

<font color = "orange">Why do we care about race? </font>Malheur County is significantly more diverse than the state of Oregon as a whole. When the topic of race came up, stakeholders noted the change as a positive. However, as a community becomes more diverse, the resources that once made sense for that community might need to be reconsidered. For example, an increasing number of Latinx and AAPI residents in a county could require increased language services that might not have existed before. Acknowledging the diversity of a population allows for more targeted policies that support these communities, leading to better economic outcomes for all residents. 


See the population map for more.



<div id="racemap">
```{r, echo=F, fig.height=6,fig.width=10}
demoChgMap

```
</div>
#
#

<h6>Data Source: U.S. Census Bureau (2010-2018), American Community Survey, 5-year estimates.</h6>

#
#
<p style="text-align:right">
<a href="https://eobed-dspg.github.io/povertyPage.html">Economic Indicators: Poverty >> </a></p>

<p style="text-align:right">
<a href="https://eobed-dspg.github.io/genPage,html"> << Back to General Main </a></p>