---
title: "Licensure"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<center>
**THIS IS A DRAFT. DO NOT DISTRIBUTE. DO NOT CITE WITHOUT AUTHOR'S PERMISSION**
</center>

#
#

One of the challenges that border counties face is competition for workers. A person can choose to be a teacher in Oregon or move two miles east and work in Idaho. A barrier that workers might face is state-specific licensing requirements. Idaho and Oregon currently have varying requirements for many occupations, thereby restricting an individual’s ability to transfer their skill set between the two states. Stakeholders shared the general sentiment that too few occupations offer provisional licenses. When provisional licenses are available, they are too challenging to acquire. This can be a driving factor in whether or not newcomers to the border area choose to live and work in Oregon rather than Idaho.


Using median annual wage data in non-metropolitan regions (the Eastern Oregon region and the Southwestern Idaho region) from the Bureau of Labor Statistics, we find that workers in certain fields often earn more in Oregon than in Idaho. This is also true for jobs that Oregon based establishments have difficulty filling and supports claims made by stakeholders. For example, the median annual wage for a registered nurse in 2018 was $77,090 in Oregon, 20 percent higher than the median wage of $64,180 in Idaho. Similarly, electricians earned a median wage of $54,390 in Idaho compared to $68,520 in Oregon. Despite the possibility of earning higher wages, individuals face barriers that make obtaining employment in Oregon difficult. For example, while an electrician can earn more in Oregon, they would need to gain more experience, meet higher continuing education requirements, and pay more to acquire their license and keep it active. Registered nurses, nursing assistants, physical therapists, teachers, and contractors face similar barriers. 


While we know that licensing barriers exist, publicly available data does not tell us much about the direct role they play in an individual’s location choice. To understand if differences in licensing requirements are responsible for driving the choice of where to live and work, we would need additional information about individual’s jobs, salaries, housing, and family, among other things. Stakeholders are encouraged to consider collecting this type of data to understand more about the role of licensing.


In the final report, we summarize the existing research on this topic for stakeholders, provide more details on wages across the two regions, and provide a summary of the National Conference of State Legislatures licensing requirement table. 


```{r, include=F}
df <- read.csv(file="../Licensure/Licensure_Table_2020.csv")
df <- df[2:length(df),]
```

```{r, echo=F}
knitr::kable(df, format = "pipe", padding = 0)
```