---
title: "Population"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=F}
source("../Population_Map/population_Map.R", local = knitr::knit_global())

```
<center>
**THIS IS A DRAFT. DO NOT DISTRIBUTE. DO NOT CITE WITHOUT AUTHOR'S PERMISSION**
</center>

#
#

<a href="#popmap">Jump to the map</a>

The population in Malheur county is 30,431 (United States Census Bureau, American Community Survey, 2018), 1.3 times greater than the population of Payette County (23,401), and three times the size of Washington County (10,025). Canyon County has a significantly larger population at 212,230, attracting families who choose to live closer to the Boise metro area. While the population in  Washington and Payette County has slightly increased in recent years, Malheur County has seen a slight decline in its overall population between 2010 and 2018. The Census tract that includes the city of Ontario saw a population decline of 6 percent between 2010 and 2018, while the Census tract that includes Fruitland saw an increase of 4 percent. The tracts just outside Ontario, on the other hand, saw a population increase of 4 percent and 6 percent. Anectodely, stakeholders believe that preferences for ranchette homes and larger properties are responsible for the growth in less populated areas outside Ontario. They also noted that the Fruitland area provides more options for those looking to purchase a home. 

<div id="popmap">

```{r, echo=F, fig.height=6,fig.width=10}
populationMap

```
</div>

#
#
Data Source: U.S. Census Bureau (2010-2018), American Community Survey, 5-year estimates.

#
#
<p style="text-align:right">
<a href="https://eobed-dspg.github.io/demoChgPage.html">Race >> </a></p>

<p style="text-align:right">
<a href="https://eobed-dspg.github.io/genPage,html"> << Back to General Main </a></p>